Making Q-learning robust to time discretization

### 2019-01-17 paper/icml_drau/framework.tex
Commit

  Many reinforcement learning environments are not intrinsically time
	discrete, but discretization of an underlying continuous time
	environment. For instance, many simulated control environments, such as the
	Mujoco environments~\cite{ddpg} or OpenAI Gym classic\TODO{CITE!} control
	environments are discretization of continuous time control problems,
        with a discretization timestep $\deltat$ fixed by the simulator.
	For real world environments, the discretization timestep, $\deltat$ is
	most often imposed by the quality of sensors and actuators. As $\deltat$ is not an
	intrinsic parameter of the environment, but a constraint imposed by
	computational ressources, we would expect that decreasing $\deltat$
	could  only improve the performance of reinforcement learning agents.
	To achieve this goal, reinforcement learning algorithms should be
	resilient to a change of $\deltat$, and should remain viable when
	$\deltat \rightarrow 0$. We are thus interested in the behavior of
	reinforcement learning algorithms in discretized environments, when the
	discretization timestep is small. We will refer to such environments as
	near continuous environments.

	In what follows, a formalized view of near continuous environments is given,
	along with $\deltat$ dependant definitions of return, discount factor and
	value functions that are shown to converge to well defined continous time limits. Finally, it is shown
	that the state-action value function collapses to the value function as
	$\deltat$ goes to $0$. Consequently there is no Q-learning in
	continuous time, foreshadowing problematic behaviors of near continuous time Q-learning.
      \TODO{We should explain somwhere here that, if not invariant, harder to find good hyperparameters}
\subsection{Framework}

@@

        Let ${\cal S} = \mathbb{R}^d$ be a set of states, and ${\cal A}$ be a set of
	actions. Consider the continuous \emph{Markov Decision Process} (MDP) defined by the \emph{Stochastic Differential Equation} (SDE)\TODO{Explain the intuition of Brownian in SDE}:
        \begin{equation}
		\label{eq:diffusion}
		ds = F(s, a)dt + \Sigma(s, a)\dbt.
              \end{equation}
        For any $\deltat$, the discrete MDP ${\cal
	M}_\deltat = \langle {\cal S}, {\cal A}, T_{\deltat}, r_\deltat,
      \gamma_\deltat\rangle$ is a discretization of the continuous MDP with \emph{time discretization} $\deltat$. The transition probability $T_\deltat(s, a, s')$ is the probability distribution of $s_\deltat$, when starting at $s_0 = s$ and maintaining $a_t=a$ constant for a time $\deltat$.
      \footnote{Informally, $T_\deltat(s, a, s')$ is approximately the probability distribution of $s + F(s, a)\deltat + \sqrt{\deltat}\Sigma(s, a){\cal N}(0, I)$, which is the discretization of Eq. \eqref{eq:diffusion} \TODO{CITE}}. This corresponds to an agent evolving in the continuous environment described by Eq. \eqref{eq:diffusion}, but who is only making observations and choosing actions every $\deltat$. We say the the MDP ${\cal M}_\deltat$ is near-continuous.

      Informally $F: {\cal S}\times{\cal A}\rightarrow {\cal S}$ describes
	the deterministic dynamic of the environment, while $\Sigma(s, a)\dbt$
	describes its stochasticity. Notably, $\Sigma = 0$ if the environment is
	deterministic, as in classical control theory, or in many usual simulated
	environments.

        In this paper, we try to design RL methods for near continuous MDP
	robust to the choice of $\deltat$. For robustness, a necessary condition is to remain viable when $\deltat \rightarrow 0$. Our method is thus to analyse the different
	quantities involved when $\deltat \rightarrow 0$, and make sure that they converge to meaningful limits.

        The first step in that direction is to check that the discretized
          trajectories converge to the continuous ones.

	\begin{theorem}
		Assume that $\pi: {\cal S} \rightarrow {\cal A}$ is a continuous and
		deterministic policy. Let $(s_t)_{t\in\mathbb{R}_+}$ be the
		continuous process by following \eqref{eq:diffusion} and
		$(s_\deltat^k)$ be the discrete process obtained if the agent only
		get observations every $\deltat$ \TODO{explain better}.

		We define the time process $\tilde s_\deltat^t = s^{\lfloor
		\frac{t}{\deltat}\rfloor}_\deltat$ for
		$t\in\mathbb{R_+}$.\TODO{Explain}. Then:
		\begin{equation}
			\label{eq:conv-traj}
			(\tilde s_\deltat^t)_{t\in\mathbb{R}_+} 
			\xrightarrow[\deltat \rightarrow 0]{} (s_t)_{t\in\mathbb{R}_+}
		\end{equation}
		in distribution.
	\end{theorem}

>>>>

  Many reinforcement learning environments are not intrinsically time
	discrete, but discretization of an underlying continuous time
	environment. For instance, many simulated control environments, such as the
	Mujoco environments~\cite{ddpg} or OpenAI Gym classic control
	environments~\cite{gym} are discretization of continuous time control problems.
	In simulated environments, the time discretization is fixed by the simulator,
	and is often used to approximate an underlying time continuous differential equation.
	It can, for example, correspond to the inverse of the number of frames generated by second.
	For real world environments, sensors and actuators have a fixed time precision: cameras can
	only capture a fixed amount of frames per second, and physical limitations prevent actuators
	from responding instantaneously. The quality of these components thus impose
	a lower bound on the discretization timestep. As the time discretization 
	$\deltat$ is not an intrinsic parameter of the environment, but a constraint imposed by
	computational ressources, we would expect that decreasing $\deltat$
	could  only improve the performance of reinforcement learning agents.
	This implies that reinforcement learning algorithms should, at least, be
	resilient to a change of $\deltat$, and should remain viable when
	$\deltat \rightarrow 0$. Besides, designing a time discretization invariant algorithm
	should alleviate tedious hyperparameterization by providing better defaults
	for time horizon related parameters.
	We are thus interested in the behavior of
	reinforcement learning algorithms in discretized environments, when the
	discretization timestep is small. We will refer to such environments as
	near continuous environments.

	In what follows, a formalized view of near continuous environments is given,
	along with $\deltat$ dependant definitions of return, discount factor and
	value functions that are shown to converge to well defined continous time limits. Finally, it is shown
	that the state-action value function collapses to the value function as
	$\deltat$ goes to $0$. Consequently there is no Q-learning in
	continuous time, foreshadowing problematic behaviors of near continuous time Q-learning.
%      \TODO{We should explain somwhere here that, if not invariant, harder to find good hyperparameters} DONE
\subsection{Framework}

@@

% TODO: For now, differential equation is the default, SDE is a footnote.
        Let ${\cal S} = \mathbb{R}^d$ be a set of states, and ${\cal A}$ be a set of
	actions. Consider the continuous \emph{Markov Decision Process} (MDP) defined by the \emph{Differential Equation}: %\TODO{Explain the intuition of Brownian in SDE}:
        \begin{equation}
		\label{eq:diffusion}
		ds = F(s, a)dt. % + \Sigma(s, a)\dbt.
              \end{equation}
        For any $\deltat$, the discrete MDP ${\cal
	M}_\deltat = \langle {\cal S}, {\cal A}, T_{\deltat}, r_\deltat,
      \gamma_\deltat\rangle$ is a discretization of the continuous MDP with
      \emph{time discretization} $\deltat$. The transition probability
      $T_\deltat(s, a, s')$ is a dirac on $s_\deltat$, where $s_\deltat$ is the
      state obtained
      when starting at $s_0 = s$ and maintaining $a_t=a$ constant for a time
      $\deltat$.
      This corresponds to an agent evolving in the continuous
      environment described by Eq. \eqref{eq:diffusion}, but who is
      only making observations and choosing actions every $\deltat$. We
      say the the MDP ${\cal M}_\deltat$ is near-continuous.\footnote{
	      We believe the results presented here to be true in a more general framework, where states
	      follow a \emph{Stochastic Differential Equation} $ds = F(s, a)dt.  + \Sigma(s, a)\dbt$, with
	      $B_t$ a multidimensional brownian motion and $\Sigma$ its associated covariance matrix. However,
      formal treatment of SDEs is generally tedious and out of the scope of this paper.}
      Informally $F: {\cal S}\times{\cal A}\rightarrow {\cal S}$ describes
	the dynamic of the environment. The agent interacts with the environment through a deterministic policy function
	$\pi \colon {\cal S} \rightarrow {\cal A}$. Determinism of the policy is assumed to prevent problematic definitions
	of the value function in continuous time. This is only a mild requirement, as we introduce an off-policy algorithm,
	and only impose determinism of the exploitation policy.
	% This framework covers classical control theory, as well as many usual simulated environments. TODO: not convinced that this is relevant here.

%      \footnote{Informally, $T_\deltat(s, a, s')$ is approximately
%	      the probability distribution of $s + F(s, a)\deltat +
%	      \sqrt{\deltat}\Sigma(s, a){\cal N}(0, I)$, which is the
%	      discretization of Eq. \eqref{eq:diffusion} \TODO{CITE}}.
        In this paper, we try to design RL methods for near continuous MDP
	robust to the choice of $\deltat$. For robustness, a necessary condition is to remain viable when $\deltat \rightarrow 0$. Our method is thus to analyse the different
	quantities involved when $\deltat \rightarrow 0$, and make sure that they converge to meaningful limits.

        The first step in that direction is to check that the discretized
          trajectories converge to the continuous ones. 
	\begin{theorem}
		Assume that $\pi: {\cal S} \rightarrow {\cal A}$ is a continuous and
		deterministic policy. Let $(s_t)_{t\in\mathbb{R}_+}$ be the
		continuous process obtained by following \eqref{eq:diffusion} and
		$(s_\deltat^k)$ be the discrete process obtained if the agent only
		get observations and interacts with the environment every $\deltat$ \TODO{explain better}.

		We define the time process $\tilde s_\deltat^t = s^{\lfloor
		\frac{t}{\deltat}\rfloor}_\deltat$ for
		$t\in\mathbb{R_+}$.\TODO{Explain}. Then:
		\begin{equation}
			\label{eq:conv-traj}
			(\tilde s_\deltat^t)_{t\in\mathbb{R}_+} 
			\xrightarrow[\deltat \rightarrow 0]{} (s_t)_{t\in\mathbb{R}_+}.
		\end{equation}
	\end{theorem}

### 2019-01-17 paper/icml_drau/intro.tex
Commit

In this paper we focus our attention on the sensitivity to time discretization
of DRL approaches,
i.e. what happens when an agent receives $50$ observations
and is expected to take $50$ actions per second instead of $10$. One expects
that decreasing time discretization, or equivalently shortening reaction time,
should only improve agent performance. Robustness to time discretization is notably relevant in \emph{near-continuous}
environments, which includes most continuous control environments, robotics and many video games.
More generally, robusteness to time discretization is important in any setting in which the advantage
an individual action over another is small compared to the total return, and notably on long time horizon
problems.

>>>>

In this paper we focus our attention on the sensitivity to time discretization
of DRL approaches,
i.e. what happens when an agent receives $50$ observations
and is expected to take $50$ actions per second instead of $10$. One expects
that decreasing time discretization, or equivalently shortening reaction time,
should only improve agent performance. Robustness to time discretization is notably relevant in \emph{near-continuous}
environments, which includes most continuous control environments, robotics and many video games.
More generally, robusteness to time discretization is important in any setting in which the advantage
of an individual action over another is small compared to the total return, and notably on long time horizon
problems.
